---
title: "Solutions_Ramon_Schildknecht_RTP Excercise Sheet - Series 2"
output:
  html_document:
    df_print: paged
---

```{r message=FALSE, warning=FALSE}
library(tseries)
library(tidyverse)
library(plotly)
library(timetk)
library(lubridate)
library(pracma)
library(fma)
```

# Template

a)

```{r}
# to do
```

b)

```{r}
# to do
```

c)

```{r}
# to do
```

d)

```{r}
# to do
```



# Series 2

## Exercise 2.1 - Match plots to correlograms

Assuming the following solutions by comparing the given pictures:
A -> 4, B -> 3, C -> 2, D -> 1 

Checking by code:

```{r}
par(mfrow=c(2,4))
# Data set cow temperature
plot(cowtemp)
acf(cowtemp)
# Data set air passengers
plot(AirPassengers)
acf(AirPassengers)
# Data set mink trappings
plot(mink)
acf(mink)
# Data set accidential deaths in the US
plot(usdeaths)
acf(usdeaths)
```

True results:
A -> 2, B -> 3, C -> 1, D -> 4


## Exercise 2.2 - calculate the lagged scatter plot and the plug-in estimator
skip for the moment




## Exercise 2.3 - AR(1) process

a)

```{r}
d <- arima.sim(list(ar = c(0.8)), n = 1000)
plot(d)
```

b)

```{r}
par(mfrow=c(2,1))
plot(ARMAacf(ar = c(1), lag.max = 100))
acf(d, lag.max = 100)
```

c)

```{r}
# assumption k = 100 as before
plot(ARMAacf(ar = c(0.8), lag.max = 100), ylim = c(-1, 1))
```

Dependence is decreasing fast from Lag 1 to lag 20. Afterwards the dependence continues stable towards zero. 

d)

```{r}
par(mfrow=c(2,1))
plot(ARMAacf(ar = c(1), lag.max = 100, pacf = TRUE))
pacf(d, lag.max = 100)
```

Observation: good agreement only for the first lag. Theoretical PACF just generates a value for lag 1. All other values are NAs!


## Exercise 2.4

a)

```{r}
# to do
```

b)

```{r}
# to do
```

## Exercise 2.5

a)

```{r}
# to do
```

b)

```{r}
# to do
```

c)

```{r}
# to do
```

d)

```{r}
# to do
```

e)

```{r}
# to do
```


## Exercise 2.6

a)

```{r}
# to do
```

b)

```{r}
# to do
```


## Exercise 2.7
a)

```{r}
# to do
```

b)

```{r}
# to do
```

c)

```{r}
# to do
```