---
title: "Solutions_Ramon_Schildknecht_RTP Excercise Sheet - Series 4"
author: "[Ramon Schildknecht](https://www.linkedin.com/in/ramon-schildknecht-36301756/)"
output:
  html_document:
    df_print: paged
    toc: true
    theme: united
  word_document: default
---

<br>
<br>

```{r packages, message=FALSE, warning=FALSE, paged.print=FALSE}
library(tseries)
library(tidyverse)
library(magrittr)
library(plotly)
library(timetk)
library(lubridate)
library(pracma)
library(fma)
library(forecast)
install.packages("fpp")
library(fpp)
```

<br>
<br>

# Series 4

<br>

## Exercise 4.1 - simulations

a)

ARMA(1,2) model

```{r}
# AR(1,2) model - theoretical autocorrelations
plot(0:30, ARMAacf(ar = c(0.75), ma = c(-0.3, 0.25), lag.max = 30), type = "h",
ylab = "ACF")

## Theoretical partial autocorrelations
plot(1:30, ARMAacf(ar = c(0.75), ma = c(-0.3, 0.25 ), lag.max = 30, pacf = TRUE), type = "h", ylab = "PACF")

```


b)

ARMA(2,1) model

```{r}
# AR(2,1) model - theoretical autocorrelations
plot(0:30, ARMAacf(ar = c(0.75, -0.3), ma = c(0.25), lag.max = 30), type = "h",
ylab = "ACF")

## Theoretical partial autocorrelations
plot(1:30, ARMAacf(ar = c(0.75, -0.3), ma = c(0.25), lag.max = 30, pacf = TRUE), type = "h", ylab = "PACF")

```


c)

```{r}
ts.sim <- arima.sim(list(order = c(2,1,2), ar = c(0.5, 0.5), ma = c(-0.4, 0.3)), n = 200)
```

Because AR part of model is not stationary!

d)

ARMA(3,1) is the equivalent model because we can take the differencing term (1 -B)^d and use it in the AR part. 

<br>


## Exercise 4.2 - sunspotarea

a)

```{r}
d <- sunspotarea
glimpse(d)
hist(d)
```

Because it is right-skewed.


b)

```{r}
d2 <- d %>% log()
d_to_1974 <- window(d, end = 1974)
tsdisplay(d_to_1974)
acf(d_to_1974)
pacf(d_to_1974)
(fit <- arima(d_to_1974, order = c(10, 0, 3)))
tsdisplay(fit$residuals)
```

Suitable model: ARMA(10, 3). Residuals looking good.


<br>



## Exercise 4.3 - beluga whales model

a)

```{r}
d_beluga <- read.table("http://stat.ethz.ch/Teaching/Datasets/WBL/beluga.dat", header = TRUE)
d_beluga_ts <-  ts(d_beluga)
(fit <- lm(NURSING ~ ., d_beluga))
summary(fit)
```

Bouts and lock-ons seems to have a significant impact on nursing success. Day or night is not important. 

b)  hier weiter

```{r}

```


c)

```{r}

```


d)

```{r}

```

e)

```{r}

```


<br>


## Exercise 4.3 - varvas

a)

```{r}

```


b)

```{r}

```


c)

```{r}

```


d)

```{r}

```

<br>




