---
title: "Solutions_Ramon_Schildknecht_RTP Excercise Sheet - Series 4"
author: "[Ramon Schildknecht](https://www.linkedin.com/in/ramon-schildknecht-36301756/)"
output:
  html_document:
    df_print: paged
    toc: true
    theme: united
  word_document: default
---

<br>
<br>

```{r packages, message=FALSE, warning=FALSE, paged.print=FALSE}
library(tseries)
library(tidyverse)
library(magrittr)
library(plotly)
library(timetk)
library(lubridate)
library(pracma)
library(fma)
library(forecast)
library(fpp)
library(car)
library(astsa)
library(dygraphs)
```

<br>
<br>

# Series 5

<br>

## Exercise 5.1 - Stationarity Rederal Reserve Board Production Index

a) 

```{r}
d <- prodn
plot(d)
```

There is a clear trend recognisable.


b)

There are several available options:

- Linear regression
- Backward difference
- Filtration

c)

First guess: SARIMA could fit.

```{r}
d_decomposition <- stl(d, s.window = 12)
plot(d_decomposition)
d_dygraph <- cbind(origin = d,
                   seasonality = d_decomposition$time.series[,"seasonal"], 
                   trend = d_decomposition$time.series[,"trend"], 
                   remainder = d_decomposition$time.series[,"remainder"])
dygraph(d_dygraph)
# dygraph(d_dygraph[,"trend"])
tsdisplay(d)
(d_sarima <- arima(d, order = c(1,1,1), seasonal = c(1, 1, 1)))
(d_sarima <- arima(d, order = c(1,1,0), seasonal = c(1, 1, 1)))
(d_sarima <- arima(d, order = c(1,0,1), seasonal = c(1, 1, 1)))
(d_sarima <- arima(d, order = c(1,1,1), seasonal = c(0, 1, 1))) # improved aic
(d_sarima <- arima(d, order = c(1,1,1), seasonal = c(0, 0, 1))) 
(d_sarima <- arima(d, order = c(1,1,1), seasonal = c(0, 2, 0)))
(d_sarima <- arima(d, order = c(1,1,1), seasonal = c(0, 1, 3))) # improved aic
(d_sarima <- arima(d, order = c(1,1,1), seasonal = c(0, 1, 4)))
```

The best result performs the following SARIMA model: <br>
arima(d, order = c(1,1,1), seasonal = c(0, 1, 3)))


<br>


## Exercise 5.2 - sunspot data & AR model

Preparation

```{r}
lsunspot100 <- window(log(sunspotarea), start = 1875, end = 1974)
fit_ar10 <- arima(lsunspot100, order = c(10, 0, 0))
```

<br>

a)

```{r}
lsunspot100 <- window(log(sunspotarea), start = 1875, end = 1974)
fit_ar10 <- arima(lsunspot100, order = c(10, 0, 0))
predictions <- predict(fit_ar10, n.ahead = 100)
plot(lsunspot100, xlim = c(1875, 2074), ylab = "yearly avg sunspot area")
abline(h = fit_ar10$coef["intercept"], lty = 2, col = "blue")
lines(predictions$pred, col = "green")
```

The predictions show a damped sinusoid towards the global mean. Every cycle reduces the scattering of the predicted points. 

b)

```{r}
lsunspot100_2 <- window(log(sunspotarea), start = 1875, end = 2011)
fit_ar10_2 <- arima(ts(lsunspot100_2, start = 1875, end = 1974), order = c(10, 0, 0))
predictions_2 <- predict(fit_ar10_2, n.ahead = 37)
plot(lsunspot100_2, xlim = c(1875, 2074), ylab = "yearly avg sunspot area")
abline(h = fit_ar10_2$coef["intercept"], lty = 2, col = "blue")
lines(predictions_2$pred, col = "green")

cat("Mean squared forecasting error: ", (msfe <- (sum((as.numeric(tail(lsunspot100_2, 37)) - as.numeric(predictions_2$pred))^2))/length(lsunspot100_2)))
plot((as.numeric(tail(lsunspot100_2, 37)) - as.numeric(predictions_2$pred))^2, type = "l",  xlab = "time", ylab = "msfe")

```

The predictions getting more inaccurate over time The mean squared forecasting error (msfe) is about 0.2. The msfe is very high at the end because there is a outlier in the time series. One could expect that the model will not fit very well for future predicted points.

<br>



## Exercise 5.3 - SARIMA with airplane data

data preparation

```{r}
d <- AirPassengers
d_1949_to_1956 <- window(d, end = c(1956, 12))
plot(d_1949_to_1956)
```


a) 

```{r}
hist(d_1949_to_1956)
plot(stl(d_1949_to_1956, s.window = 12))
tsdisplay(d_1949_to_1956)
tsdisplay(diff(d_1949_to_1956))
pacf(d_1949_to_1956) # lag 11 seems suitable for ar()
d_1949_to_1956_arima_coefficients <- auto.arima(d_1949_to_1956)
fit_1949_to_1956 <- arima(d_1949_to_1956, order = c(1, 1, 0), seasonal = c(1, 1, 0))
predictions <- predict(fit_1949_to_1956, n.ahead = 48)

plot(d_1949_to_1956, lty = 3, xlim = c(1949, 1961), ylim = c(0, 800), ylab = "air passengers (unit = 1000)")
lines(d_1949_to_1956, lwd = 1)
lines(predictions$pred, lwd = 2, col = "darkred")
lines(predictions$pred+predictions$se*1.96, col="tomato")
lines(predictions$pred-predictions$se*1.96, col="tomato")
lines(window(d, start = c(1957, 01), lwd = 2, end = c(1960, 12)), col = "springgreen")


# the easier way with forecast | dark grey = 80% confidence interval | grey = 90% ci)
plot(forecast(fit_1949_to_1956, h = 48))
lines(window(d, start = c(1957, 01), lwd = 2, end = c(1960, 12)), col = "springgreen")
```



b)


c)


<br>
