---
title: "Solutions_Ramon_Schildknecht_RTP Excercise Sheet - Series 1"
output:
  html_document:
    df_print: paged
---

```{r message=FALSE, warning=FALSE}
library(tseries)
library(tidyverse)
library(plotly)
library(timetk)
library(lubridate)
library(plotly)
library(pracma)
```


# Series 1

## Exercice 1.1 - backward differencing

```{r}
set.seed(22)
# Apply backward differencing on a timeseries created by the following model:
# create timeseries object
t <- seq(1, 100, length = 100)
data <- 2 * t + 0.5 + runif(100, -5, 5)
ts <- ts(data)
# backward differencing
ts_diff <- diff(ts)
# compare lengths of arrays with
print("length difference:")
(length_diff <- length(ts) - length(ts_diff))
# adf.test(ts) # be careful!
plot(ts)
plot(ts_diff)
```

## Exercice 1.2 - remove linear trend
a)

```{r}
# create reusable function
generate_and_plot_ts <- function(model) {
  set.seed(22)
  t <- seq(1, 100, length = 100)
  data <- model
  ts <- ts(data)
  
  # backward differencing
  ts_diff <- diff(ts)
  plot(ts)
  plot(ts_diff)
}
```


```{r}
generate_and_plot_ts(0.5*t + 1 + runif(100, -1, 1))
```


b)
```{r}
generate_and_plot_ts(2*t^2 + 3*t + runif(100, -200, 200))
```


## Exercice 1.3

```{r}
d <- co2
glimpse(d)
head(d)
tail(d)
# already a time series object!
# ts <- ts(d, start = 1959, end = 1998, frequency = 12)
plot(co2, main = "co2 data")

# generate interactive plot
d_index <- tk_index(d)
(p <- plot_ly(x = ~d_index, y = ~d, mode = 'lines'))

acf(d)
pacf(d)
# try and error
# d_diff <- (diff(d, lag = 1))
# d_diff2 <- (diff(d_diff, lag = 1))
# d_diff3 <- (diff(d_diff2, lag = 1))
# d_diff4 <- (diff(d_diff3, lag = 1))
# d_diff5 <- (diff(d_diff4, lag = 1))

# 5 differences is optimal
d_diff <- (diff(d, lag = 1, differences = 5))
acf(d_diff)
pacf(d_diff)

# check result with interactive plot
d_diff_index <- tk_index(d_diff)
(p2 <- plot_ly(x = ~d_diff_index, y = ~d_diff, mode = 'lines'))
# success!
```


## Exercice 1.4

```{r}
# help trough https://anomaly.io/seasonal-trend-decomposition-in-r/
d <- co2
d_stl <- stl(d, s.window = "periodic")
plot(d_stl)
plot(d_stl, xlim=c(1960, 1961))

# seasonal part:
d_trend <- d_stl$time.series[, 2]
d_season <- d_stl$time.series[, 1]
d_random <- d_stl$time.series[, 3]
# check if it worked correctly
head(d-(d_trend+d_season+d_random))
# success - worked!

trend_removed <- d - d_trend
month <- factor(rep(1:12, 39))
print("Season means by month")
(seasn_est <- tapply(trend_removed, month, mean, na.rm = TRUE))
```

```{r}
# decompose gives same results as stl()
d_decomposed = decompose(d, "additive")
plot(as.ts(d_decomposed$seasonal))
plot(as.ts(d_decomposed$trend))
plot(as.ts(d_decomposed$random))
plot(d_decomposed)
```


## Exercice 1.5

```{r}
# a)
d <- read.table(file = "rainDay.txt")
d$DATE <- dmy(d$DATE)
glimpse(d)

# b)
d_ts <- ts(data = d$rain, start = c(2000, 1), end = c(2007, 12), frequency = 12)

# c)
d_extended <- d %>% 
  mutate(day_of_week = as.factor(weekdays(DATE)),
         month = as.factor(months(DATE)),
         quarter = as.factor(quarters(DATE))) 
glimpse(d_extended) # worked!

hist(d_ts)
# hint: the series is right skewed - one should log transform it but...

par(mfrow=c(2,1)) 
plot(d_ts)
plot(log(d_ts))

# ... one do not log transform in this case --> otherwise we lose observations!
# additionally: the time series is not multiplicative

# d)
# set logical order for weekday and month
levels(d_extended$day_of_week) = c("Montag", "Dienstag", "Mittwoch", "Donnerstag", "Freitag", "Samstag", "Sonntag")

levels(d_extended$month) = c("Januar", "Februar", "März", "April", "Mai", "Juni", "Juli", "August", "September", "Oktober", "November", "Dezember")

(weekday_boxplot <- d_extended %>%
  # group_by(day_of_week) %>% 
  # mutate(sum_rain_daily = sum(rain)) %>% 
  # distinct(day_of_week, .keep_all = T) %>% 
  # ungroup() %>% 
  # select(day_of_week, rain) %>% 
  plot_ly(y = ~rain, color = ~day_of_week, type = "box"))

(monthly_boxplot <- d_extended %>%
  plot_ly(y = ~rain, color = ~month, type = "box"))

(quarter_boxplot <- d_extended %>%
  plot_ly(y = ~rain, color = ~quarter, type = "box"))

# e) 
(rain_plot <- plot_ly(x = ~d$DATE, y = ~d$rain, mode = 'lines'))
  
```

## Exercice 1.6

a) period: year & time step: day 
b) period: year & time step: day 
c) period: weeks & time step: hours to day 
d) period: year & time step: minutes to day  



## Exercice 1.7

```{r}
d <- read.table("hstart.dat.txt")
ts <- ts(d, start = c(1966, 1), end = c(1974, 12), frequency = 12)

# a) 
(hstart_plot <- plot_ly(x = ~time(ts) , y = ~ts, mode = 'lines'))

# b)
# skipped loess method, decompose is easier to use
d_decomposed = decompose(ts, "additive")
plot(as.ts(d_decomposed$seasonal))
plot(as.ts(d_decomposed$trend))
plot(as.ts(d_decomposed$random))
plot(d_decomposed)
```

a) + b) time series is not stationary. Points (1), (2) and (3) mentioned below in Exercice 1.8 are not met. 
One can see a clear yearly trend. One can further recognize a trend with two peaks. 
Conclusion: Get seasonality and trend component out of the time series. 

c) do not know how to code the special filter


```{r}
# d)
# remove trend
ts_no_seasonality <- ts - d_decomposed$seasonal
# ts_no_seasonality_and_no_trend <- ts_no_seasonality - d_decomposed$trend
# check results
# d_decomposed$random - ts_no_seasonality_and_no_trend

# apply linear trend elimination
install.packages("pracma")
ts_no_seasonality_no_linear_trend <- pracma::detrend(ts_no_seasonality, tt = "linear")

# plot data
(hstart_plot_2 <- plot_ly(x = ~time(ts) , y = ~ts_no_seasonality_no_linear_trend, mode = 'lines'))

```




## Exercice 1.8

Stationarity if...
- (1) all Xt are identically distributed 
- (2) regular mean 
- (3) finit and regular variance 
- (4) autocovariance dependes only on lag 


```{r warning=FALSE}
# a)
Et <- rnorm(100, 0, 1)
Et_1 <- diff(Et)
model <- Et[2:100] - (0.5 * Et_1)
ts <- ts(model)
(Y1_plot <- plot_ly(x = ~c(1:99), y = ~model, mode = 'lines'))
acf(ts)
pacf(ts)

```

a) all criterias are allright - this time series is stationary.


b - d) do not understand questions. What is Yt compared with Et?


```{r}
# b)
Et <- rnorm(100, 0, 1)
Et_1 <- diff(Et)
model <- Et[2:100] - (0.5 * Et_1)
ts <- ts(model)
(Y1_plot <- plot_ly(x = ~c(1:99), y = ~model, mode = 'lines'))
acf(ts)
pacf(ts)

```










