---
title: "Solutions_Ramon_Schildknecht_RTP Excercise Sheet - Series 4"
author: "[Ramon Schildknecht](https://www.linkedin.com/in/ramon-schildknecht-36301756/)"
output:
  html_document:
    df_print: paged
    toc: true
    theme: united
  word_document: default
---

<br>
<br>

```{r packages, message=FALSE, warning=FALSE, paged.print=FALSE}
library(tseries)
library(tidyverse)
library(magrittr)
library(plotly)
library(timetk)
library(lubridate)
library(pracma)
library(fma)
library(forecast)
library(fpp)
library(car)
```

<br>
<br>

# Series 4

<br>

## Exercise 4.1 - simulations

a)

ARMA(1,2) model

```{r}
# AR(1,2) model - theoretical autocorrelations
plot(0:30, ARMAacf(ar = c(0.75), ma = c(-0.3, 0.25), lag.max = 30), type = "h",
ylab = "ACF")

## Theoretical partial autocorrelations
plot(1:30, ARMAacf(ar = c(0.75), ma = c(-0.3, 0.25 ), lag.max = 30, pacf = TRUE), type = "h", ylab = "PACF")

```


b)

ARMA(2,1) model

```{r}
# AR(2,1) model - theoretical autocorrelations
plot(0:30, ARMAacf(ar = c(0.75, -0.3), ma = c(0.25), lag.max = 30), type = "h",
ylab = "ACF")

## Theoretical partial autocorrelations
plot(1:30, ARMAacf(ar = c(0.75, -0.3), ma = c(0.25), lag.max = 30, pacf = TRUE), type = "h", ylab = "PACF")

```


c)

```{r eval=FALSE, include=FALSE}
ts.sim <- arima.sim(list(order = c(2,1,2), ar = c(0.5, 0.5), ma = c(-0.4, 0.3)), n = 200)
```

Because AR part of model is not stationary!

d)

ARMA(3,1) is the equivalent model because we can take the differencing term (1 -B)^d and use it in the AR part. 

<br>


## Exercise 4.2 - sunspotarea

a)

```{r}
d <- sunspotarea
glimpse(d)
hist(d)
```

Because it is right-skewed.


b)

```{r}
d2 <- d %>% log()
d_to_1974 <- window(d, end = 1974)
tsdisplay(d_to_1974)
acf(d_to_1974)
pacf(d_to_1974)
(fit <- arima(d_to_1974, order = c(10, 0, 3)))
tsdisplay(fit$residuals)
```

Suitable model: ARMA(10, 3). Residuals looking good.


<br>



## Exercise 4.3 - beluga whales model

a)

```{r}
d_beluga <- read.table("http://stat.ethz.ch/Teaching/Datasets/WBL/beluga.dat", header = TRUE)
d_beluga_ts <-  ts(d_beluga)
(fit <- lm(NURSING ~ ., d_beluga))
summary(fit)
fit_res <- fit$residuals
tsdisplay(fit_res)
car::qqPlot(fit_res)
hist(fit_res)
```

Bouts and lock-ons seems to have a significant impact on nursing success. Day or night is not important. There is some structure left in the residuals. One must further investigate the dependency structure at lag 1 and 2 in (P)ACF. The residuals seem to be normally distributed.

b)  

```{r}
# use AR(2) model according to PACF result of residuals above
f_burg <- ar.burg(fit_res, order = 2)
# summary(f_burg)
f_burg$x.mean
f_burg$ar
```

The mean is close to zero (6.67e-17). The parameter alpha 1 is about 0.28 and alpha 2 about 0.32.


c)

```{r}
library(nlme) # Load the package containing the procedure gls()
d_beluga_gls <- gls(NURSING ~ BOUTS + LOCKONS + DAYNIGHT + PERIOD, 
                    data = d_beluga, 
                    correlation = corARMA(form = ~ PERIOD, p = f_burg$order, q = 0,
                                          fixed = FALSE), method = "ML")
summary(d_beluga_gls)
d_resid <- ts(resid(d_beluga_gls ))
```


d) - f) Leave this exercises out and use them as proper exam preparation.


<br>


## Exercise 4.3 - varvas

a)

```{r}
t_url <- "http://stat.ethz.ch/Teaching/Datasets/WBL/varve.dat"
d_varve <- ts(scan(t_url)[201:550])
glimpse(d_varve)
plot(d_varve)
hist(d_varve)
```

Because it is right-skewed.

b)

```{r}
d_varve_log <- log(d_varve)
tsdisplay(d_varve_log)
d_varve_diff_log <- diff(d_varve, differences = 1)
tsdisplay(d_varve_diff_log)
```

No, it is not stationary. There seems to be a trend because of the slow decay in the ACF. There could be a seasonal component due to the oscillatory behaviour as well. One could try to difference the series. But there is still a autocorrelation left at lag 1 to 5 in the PACF.

c)

```{r}
auto.arima(d_varve, seasonal = T, stationary = F, ic = "aic")
(varve_fit <- arima(d_varve, order = c(1, 1, 1), seasonal = c(1, 1, 1)))
na.omit(varve_fit$coef)
```


d)

Y(t) = 0.1402 * Y(t-1) + -0.9656 * E(t) + 0.1402 * ??? -0.956 * ???.

Need some help here. How to I get the results above in the correct formula?
Looked for help but nothing helped me to solve the task (e. g. https://stats.stackexchange.com/questions/129901/sarima-model-equation).

<br>




